Nail: A practical interface generator for data formats

### 2014-05-01 paper/osdi14/design.tex
Commit

\paragraph{Allow parser extension.}
Real-world protocols contain complicated ways of encoding data. Representing these in an
intentionally limited model such as our parser language is complicated. Therefore, Nail introduces
\emph{transformations}, which allow arbitrary code by the programmer to interact with the parser and
generator. Nail parsers and generators interact with data through an abstract stream, which allows
reading and writing of bits as re-positioning. Transformations allow the programmer to write
functions in a general purpose language that consume streams and define new temporary streams, while
also reading or writing the value of dependent fields.

@@

In the rest of this section, we describe the design on the Nail grammar language.
\input{fig-syntax}
\subsection{Basics}
A Nail parser defines both the structure of some external format and a data type to represent that
format. Parsers are constructed by combinators over simpler parsers, an approach popularised by the Parsec
framework\cite{LeijenMeijer:parsec}. We provide the most common combinators familiar from other parser
 combinator libraries, such as Parsec and Hammer\cite{hammer-parser} and extend them so they also
 describe a data type. 
\paragraph{Integers and Constraints}

Nail's fundamental parsers represent signed or unsigned with arbitrary lengths up to 64 bits, an
approach popularised by the Hammer framework.\cite{hammer-parser}  
Note that is possible to define parsers for sub-byte lengths, e.g. to parse the 4-bit
data offset within the TCP header.  In the internal representation, integer values are represented
by the nearest usable integer size, for example a 24-bit unsigned integer will be stored in an
unsigned 32-bit double word.

The grammar can also express a constraint on an integer field. The current version of Nail expresses
constraints as a set of permissible values or value ranges. Extending the Nail language and
implementation to support richer constraints languages would be relatively trivial, however we have
found that the current syntax covers permissible values within existing protocols correctly and
concisely.

@@

If two options the same string could be parser by two options, generated output for the latter
option is not necessarily understood identically by the parser. However, actual data formats are
usually not ambiguous in this sense.
For example, Figure~\ref{fig:syntax} demonstrates a simple choice
combinator.

@@

\paragraph{Rules.}
A Nail grammar file consists of rules that assign a parser to a name. Rules are written as
assignments, such as \texttt{ints = many uint8}, which defines a rule called \texttt{ints} with the
body \texttt{many uint8}.  As we will describe in Sections~\ref{s:dependent}
and~\ref{s:transforms}, rules can optionally consume parameters. 
Rules can be invoked in a Nail grammar anywhere a parser can appear. Rule invocations act as though
the body of the rule had been substituted in the code. If parameters appear, they are passed by reference.

>>>>

\paragraph{Parser Extensions.}
Real-world protocols contain complicated ways of encoding data. Representing these in an
intentionally limited model such as our parser language is complicated. Therefore, Nail introduces
\emph{transformations}, which allow arbitrary code by the programmer to interact with the parser and
generator. Nail parsers and generators interact with data through an abstract stream, which allows
reading and writing of bits as re-positioning. Transformations allow the programmer to write
functions in a general purpose language that consume streams and define new temporary streams, while
also reading or writing the value of dependent fields.

@@

In the rest of this section, we describe the design on the Nail grammar language.
\input{fig-syntax}
\subsection{Basics}
A Nail parser defines both the structure of some external format and a data type to represent that
format. Parsers are constructed by combinators over simpler parsers, an approach popularised by the Parsec
framework\cite{LeijenMeijer:parsec}. We provide the most common combinators familiar from other parser
 combinator libraries, such as Parsec and Hammer\cite{hammer-parser} and extend them so they also
 describe a data type.

Figure~\ref{fig:syntax} shows the elements of Nail grammars, which we describe in more detail below.
\paragraph{Integers and Constraints}

Nail's fundamental parsers represent signed or unsigned with arbitrary lengths up to 64 bits, an
approach popularised by the Hammer framework.\cite{hammer-parser}  
Note that is possible to define parsers for sub-byte lengths, e.g. to parse the 4-bit
data offset within the TCP header.  The internal representation rounds up to the nearest integer
type, for example a 24-bit unsigned integer will be stored in an unsigned 32-bit double word.

The grammar can also constraint the values of an integer. The current version of Nail expresses
constraints as a set of permissible values or value ranges. Extending the Nail language and
implementation to support richer constraints languages would be relatively trivial, however we have
found that the current syntax covers permissible values within existing protocols correctly and
concisely.

@@

If two options the same string could be parser by two options, generated output for the latter
option is not necessarily understood identically by the parser. However, actual data formats are
usually not ambiguous in this sense.

@@

\paragraph{Rules.}
A Nail grammar consists of rules that assign a parser to a name. Rules are written as
assignments, such as \texttt{ints = many uint8}, which defines a rule called \texttt{ints} with the
body \texttt{many uint8}.  As we will describe in Sections~\ref{s:dependent}
and~\ref{s:transforms}, rules can optionally consume parameters. 
Rules can be invoked in a Nail grammar anywhere a parser can appear. Rule invocations act as though
the body of the rule had been substituted in the code. If parameters appear, they are passed by reference.

### 2014-05-01 paper/osdi14/relwk.tex
Commit

% \XXX[Move to the intro, into its own section, or kill?]

\XXX[previous workshop paper~\cite{bangert:nail-langsec}.]

@@

%%This is supposed to go into design
% While it is possible to express short transformations on the input
% entirely as semantic actions\footnote{This is in fact the design
% rationale, to perform computation on the fly as the parser walks the
% parse tree.}, more complicated programs usually construct an internal
% representation, which contains all relevant information from the input
% in a format native to the programming language used. For example, a C
% programmer ideally wants to deal with structs and NULL-terminated arrays,
% whereas a C++ programmer might expect STL containers, a Java programmer
% interfaces, a Haskell programmer records and a LISP programmer property
% lists.  The structure of this internal representation usually resembles
% the structure of the grammar.

>>>>

@@

\paragraph{Workshop paper}
We presented an early design of Nail at the LangSec workshop~\cite{bangert:nail-langsec} that had
no support for transformations, as described in Section~\ref{s:transforms}, and much more limited
dependent fields, as described in Section~\ref{s:dependent}. With our extended and refined design,
Nail can support much more complicated, real-world data formats.
%%This is supposed to go into design
% While it is possible to express short transformations on the input
% entirely as semantic actions\footnote{This is in fact the design
% rationale, to perform computation on the fly as the parser walks the
% parse tree.}, more complicated programs usually construct an internal
% representation, which contains all relevant information from the input
% in a format native to the programming language used. For example, a C
% programmer ideally wants to deal with structs and NULL-terminated arrays,
% whereas a C++ programmer might expect STL containers, a Java programmer
% interfaces, a Haskell programmer records and a LISP programmer property
% lists.  The structure of this internal representation usually resembles
% the structure of the grammar.
